---
title: "Assignment 1"
author: David Lewis
output: pdf_document
---

## import the data
```{r}
Turtles <- data.frame(readxl::read_excel("PaintedTurtles.xlsx"))
```

## 1. What is the dimension of the data?
```{r}
dim(Turtles)
```

The data has 24 rows and 3 columns (three dimensions, 24 points).

## 2. Exhibit the entire data
```{r}
Turtles
```

## 3. Find the summary of each column of the data
```{r}
summary(Turtles)
```

## 4. Find the variance covariance matrix
```{r}
var(Turtles)
```

## 5. Transform the data into natural logarithm and reproduce the transformed data.

```{r}
Turtles1 <- data.frame(L1 = log(Turtles$Length), W1 = log(Turtles$Width), H1 = log(Turtles$Height))
Turtles1
```


## 6. Obtain the variance covariance of the transformed data     

```{r}
var(Turtles1)
```


## 7. Total amount of variation present in the transformed data
Sum of variance values (diagonal in variance covariance matrix).
```{r}
sum(diag(var(Turtles1)))
```

## 8. Principal component analysis on the transformed data
```{r}
Turtles2 <- princomp(Turtles1)
Turtles_prcomp <- prcomp(Turtles1)
```

## 9. Obtain the eigenvalues and eigenvectors of the variance-covariance matrix of the data. 

```{r}
eigens <- eigen(var(Turtles1))
eigens
```

## 10. Spell out the three principal components of the data.
```{r}
eigens
summary(Turtles2)
```

[, 1] = Comp.1

## 11. Calculate the percentage of variation accounted by the first principal component, first two principal components, and all principal components. 

This is taken from the summary in the above question.

- Percentage of first component: 96.05%
- Percentage of first 2 components: 98.52%
- Percentage of all components: 100%

## 12. Plot the variances of the principal component scores.
```{r}
plot(Turtles2)
```


## 13. Tally the variances with the eigenvalues of the transformed variance covariance matrix
```{r}
sum(eigens$values)
```


## 14. Calculate the mean of each of the principle component scores

```{r}
apply(Turtles2$scores, 2, mean)
```


## 15. Write the first prinicple component in terms of the original measurements

```{r}
Turtles_prcomp$rotation[, 1]
```