---
title: Homework 3
author: David Lewis
date: March 11, 2025
output:
    pdf_document: default
    # pdf_document: default
    #     template

# header-includes:
#     - \usepackage{fvextra}
#     - \DefineVerbatimEnvironment{Highlighting}{Verbatim}{breaklines,commandchars=\\\{\}}
documentclass: assignments

---
<!--
A preamble: I have analyzed the biomarker CA125 protein. It is your turn to make
something out of the protein CA19.9.
-->

## Import the data
```{r}
library("printr")
data <- read.csv("PancreaticCancer.csv")
```

## 1. Dig up some information on these proteins. 1 + 1 points
CA 19-9 is a biomarker associated with cancers relating to the colon, stomach and bile duct.
It is in fact, not a protein at all, but a tetrasaccharide known to play a role in cell-to-cell recognition processes.

CA125 on the other hand is a glycoprotein that can carry CA 19-9. As such, 
they should have similar levels of activity in cancerous cells.




## 2. I recommend using the log transform of the protein CA19.9 as the biomarker.
### Justification Function
```{r}
justification <- function(v) {
    data.frame(mean = mean(v), sd = sd(v), CV = mean(v) / sd(v))
}
```
### Log transform
```{r}
justification(data[["CA19.9"]])
justification(log(data[["CA19.9"]]))
data <- within(data, {
    CA19.9 <- log(CA19.9)
})
```

The standard deviation and coefficient of variation significantly decrease after log transformation making the data much easier to work with and visualize.

## 3. Segregate the data by cancer and pancreatitis. Obtain the kernel densities of logCA19.9 in the same graph. Make the graph as informative as possible. Comment. 1 + 3 + 1 points

### Segregate data
```{r}
cancer <- subset(data, data$Outcome == 1)
pancreatitis <- subset(data, data$Outcome == 0)
```

### Plot kernel densities
```{r}
cancer_density <- density(cancer$CA19.9)
pancreatitis_density <- density(pancreatitis$CA19.9)

ylim <- max(c(cancer_density$y, pancreatitis_density$y))

plot(density(cancer$CA19.9), xlab = "Carbohydrate Antigen 19-9", ylab = "Density", col = "red", lwd = 2, main = "Kernel Densities", ylim = c(0, ylim + 0.05))
lines(density(pancreatitis$CA19.9), col = "blue", lwd = 2)
legend("topright", legend = c("Pancreatic Cancer", "Pancreatitis"), pch = c(16, 16), col = c("red", "blue"))
rug(cancer$CA19.9, side = 1, col = "red")
rug(pancreatitis$CA19.9, side = 3, col = "blue")
```


## 4. Obtain the summary statistics of logCA19.9 by Cancer and Pancreatitis.
```{r}
summary(cancer$CA19.9)
summary(pancreatitis$CA19.9)
```

The cancer data has much more variability when compared to the pancreatitis data set. 

## 5. Lay down the structure of a potential diagnostic test. 2 points

Using the first quartile of the cancer patients as a cutoff, one can make a "classifier" for positive and negative for cancer.

- cancer positive if logCA19.9 $\ge$ 1.872
- cancer negative if logCA19.9 $<$ 1.872

### diagnostic test R implementation
```{r}
diagnostic_test <- function(x, y, cutoff = 1.872) {
    pred_x <- x >= cutoff
    TP <- sum((pred_x == y) & (pred_x == 1))
    FP <- sum((pred_x != y) & (pred_x == 1))
    TN <- sum((pred_x == y) & (pred_x == 0))
    FN <- sum((pred_x != y) & (pred_x == 0))
    data.frame(Cancer = c(TP, FN), "No Cancer" = c(FP, TN), row.names = c("Positive", "Negative"))
}
diagnostic_test(data$CA19.9, data$Outcome)
```


## 6. Big step: Present the ROC table consisting of sensitivities, and specificities for a vast range of cutoffs. 5 points
```{r}
roc <- pROC::roc(data$Outcome, data$CA19.9, levels = c(0, 1), direction = "<")
data.frame(Sensitivity = roc$sensitivities, Specificity = roc$specificities, Cutoff = roc$thresholds)
```


## 7. Present the ROC curve associated with the biomarker. Make the graph as informative as possible. 3 points
```{r}
plot(roc, col = "red", main = "ROC curve for Pancreatic Cancer of CA19.9")
```
## 8. Report the area under the curve and 95% confidence interval for the
population AUC. 1 + 1 points

```{r}
roc$auc
pROC::ci.auc(roc)
```
